% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{algorithm,caption,algpseudocode}
\usepackage{xcolor}

\newcommand\alert[1]{\textcolor{red}{#1}}
\newcommand\note[1]{\textcolor{blue}{#1}}
\newcommand\jb[1]{\textcolor{green!50!black}{#1}}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{ASSESS}{2014 New York City, USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Item Response Theory VS Q-matrices for Adaptive Testing}
%\subtitle{English subtitles
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{5} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor Jill-Jênn Vie\\
       \affaddr{ENS Cachan -- Bât. Cournot}\\
       \affaddr{61 av. du Président Wilson}\\
       \affaddr{94235 Cachan, France}\\
       \email{vie@jill-jenn.net}
% 2nd. author
\alignauthor
Fabrice Popineau\\
       \affaddr{Supélec -- Dép. informatique}\\
       \affaddr{3 rue Joliot Curie}\\
       \affaddr{91192 Gif-sur-Yvette, France}\\
       \email{fabrice.popineau@supelec.fr}
% 3rd. author
\alignauthor
Jean-Bastien Grill\\
       \affaddr{École normale supérieure}\\
       \affaddr{45 rue d'Ulm}\\
       \affaddr{75005 Paris}\\
       \email{grill@clipper.ens.fr}
\and
% 4th. author
\alignauthor Éric Bruillard\\
       \affaddr{ENS Cachan -- Bât. Cournot}\\
       \affaddr{61 av. du Président Wilson}\\
       \affaddr{94235 Cachan, France}\\
       \email{eric.bruillard@ens-cachan.fr}
% 5th. author
\alignauthor
Yolaine Bourda\\
       \affaddr{Supélec -- Dép. informatique}\\
       \affaddr{3 rue Joliot Curie}\\
       \affaddr{91192 Gif-sur-Yvette, France}\\
       \email{yolaine.bourda@supelec.fr}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
% \additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
% email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
% (The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{May 26, 2014}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Computerized Adaptive Testing (CAT) is a mode of testing which has gained increasing popularity over the past years (economic purposes): it selects the questions asked to the examinee in order to value her level very effectively, by using her answers to the previous questions.
Traditionally, CAT systems have been relying on Item Response Theory (IRT) in order to provide an effective measure of a latent ability in large-scale assessments.
In recent work in the field of psychometrics, cognitive diagnosis models have emerged, incorporating examinee skills in order to provide useful feedback. Those have been extensively studied in the literature and mostly rely on q-matrices.
In this paper, we propose a new framework for evaluating adaptive testing algorithms and use it to compare both the latent trait model using IRT and the cognitive diagnosis model using q-matrices.
We found that q-matrices perform better on both simulated data and real data.
\end{abstract}

% A category with the (minimum) three required fields
\category{}{Student assessment}{Adaptive testing}
%A category including the fourth, optional field follows...
\category{}{Active learning}{Generalized binary search}

\terms{Adaptive tests, q-matrices}

\keywords{Adaptive assessment}

\section{Introduction}
Automated assessment has been developed with the help of online initiatives such as the GMAT~\cite{Rudner2010}, or more recently with the introduction of massive online open courses (MOOCs). The fact of ranking thousands of students automatically in evaluating or recruiting purposes or providing personal feedback to thousand of students automatically in formative purposes has been of growing interest.

If we already have for a certain test a large-scale database of item responses, it is natural to wonder which questions bring the most information about an examinee, i.e. if there is a ``best'' sequence according to which the questions should be asked. In oral examinations, the examiner usually picks the next question to ask according to the past performance of the examinee. Computerized adaptive testing can be seen as an automated version of this process: keep asking the best possible questions until enough information has been gathered. Indeed, if the examinee behaves typically, a subset of carefully chosen results may be enough to guess how she will perform on the rest of the questions.

In order to address this purpose, Item Response Theory (IRT) is of great help. Although the purpose IRT was first to provide a framework for evaluating how well individual items on assessments work~\cite{Hambleton1991} by establishing a link between difficulty of items and ability of contestants, it gained interest for computerized adaptive testing and several methods were proposed and implemented. According to her performance, an examinee gets a score. The simplicity of the IRT models allowed a strong theoretical analysis~\cite{Baker2004} while increasing its popularity.

More recently, the No Child Left Behind Act of 2001 called for more formative assessments, in order to detect early students with cognitive disabilities and urged the need of cognitive diagnosis models. Instead of a summative score, examinees would receive a detailed feedback, specifying which skills are mastered and which ones are not~\cite{Cheng2009}. These models allow researchers to reveal possible misconceptions of the examinee in order to suggest her appropriate exercises. Most of these models rely on an q-matrix specifying for each question the different skills required to solve it. For example, a fraction test may require 1) converting a whole number to a fraction, 2) separating a whole number from a fraction, 3) simplifying before subtracting, and so on~\cite{DeLaTorreDouglas2004}. Recently, adaptive tests have been designed in order to guess the skills of the examinee effectively~\cite{Huebner2010}.

Originally, q-matrices were specified by experts, but recent work in the educational data mining field tend to infer such q-matrices directly from student data, using hill-climbing techniques~\cite{Barnes2005}, non-negative matrix factorization~\cite{Desmarais2011} or the EM algorithm~\cite{Huebner2010}. The skills extracted by data mining algorithms are therefore implicit, but have proven to produce better inference~\cite{Barnes2003}.

\subsection{Our contribution}

In this paper, we propose a new framework for evaluating adaptive testing algorithms and use it to compare the performances of both a traditional computerized adaptive testing using item response theory and a more recent cognitive diagnosis model from the literature in psychometrics using q-matrices. More precisely, we tend to answer the following question: given a budget of $B$ questions asked according to a certain adaptive selection rule, which model performs the best in predicting the answers of the examinee over the remaining questions?

One could wonder why we compare a model designed for assessment versus a model designed for feedback. Although q-matrix were first designed as a cognitive diagnostic model to provide a feedback, it is also possible to use them in the context of large-scale assessments. Indeed, we can estimate the performance of an examinee, according to her computed skills. Thus, the framework we propose for evaluation captures both models.

Whereas IRT is the most commonly used model for adaptive testing, Q-matrix turns out to be a better model for fine-grained cognitive diagnosis. In fact, our results show the cognitive diagnosis model outperforms IRT on both simulated and real data.

\subsection{Outline}

We will first present the computerized adaptive testing framework and two popular models chosen respectively from item response theory and cognitive diagnosis. We then detail the algorithms chosen in the two corresponding adaptive tests implementations and the proposed framework for evaluating them in our simulation. Finally, we will then present our results on both simulated and real data.

\section{Background and Related Work}

We present here two models that allow us to predict the performance of the examinee.

\subsection{Computerized adaptive testing}

In a computerized adaptive test (CAT), the examinee is presented with adaptively selected questions according to her previous performance. Thus, a CAT framework relies on two main subroutines:
\begin{itemize}
\item \textsc{NextItem} : the item selection algorithm, that will pick the next question to ask according to the previous answers of the examinee;
\item \textsc{TerminationRule} : a condition that will end the test, when enough information has been gathered and the competence has been measured satisfyingly.
\end{itemize}

Common criteria for the item selection rule are minimizing the Shannon entropy of the distribution over the parameters, or maximizing the Kullback-Leibler divergence. \alert{TODO Plus précis ?}

The framework of a CAT can be represented by the following algorithm: while the termination rule is not satisfied, the algorithm picks the question that maximizes a certain criterion according to the item selection rule and update the parameters according to the answer.

\subsection{Item Response Theory}

For our needs, IRT features a commonly used model allowing to compute the probability item response. Basically, IRT estimate the ability of a student by a unique real number $\theta$ and characterize each question by two real numbers: \alert{TODO est-ce que j'ai bien dit ce qu'était probability item response ?}

\begin{itemize}
\item the \emph{difficulty} $d$, corresponding to the required ability for answering the question correctly; \alert{TODO ok ?}
\item the \emph{discrimination} $\delta$, representing the capacity of the question to distinguish between students that have the required ability from the other ones.
\end{itemize}

Knowing the ability $\theta$ of a given student and both the difficulty $d$ and the discrimination parameter $\delta$ of a given question, the model allows us to compute the probability of the event “student answers correctly a question of difficulty $d$ and discrimination $\delta$”, there denoted by \emph{success}:
\[ \Pr\{success|\theta\} = \frac1{1+e^{-\delta(\theta - d)}} \]

The challenge is to determine the parameters that fit the most the student data. We maintain a probability distribution of $\theta$ and each question answered allows to refine the confidence interval around $\theta$ using the Bayes rule. This simple model is commonly used in computerized adaptive testing, and many methods are implemented for next item selection~\cite{MagisRaiche2012}. From the probability distribution over $\theta$, it is possible to compute the expected probability of answering every question of the bank. \alert{TODO}

\subsection{Cognitive Diagnosis Model}

Let us assume there are $K$ different skills. A user can be modeled by a vector of binary values $(a_1, \ldots, a_K)$, called \emph{state}, representing her master of the different skills. Thus, there are $2^K$ possible states.

A q-matrix \cite{Tatsuoka1983} represents, for each question, the different skills involved in answering the question. $Q_{ij}$ is equal to 1 if the skill $j$ is involved in the resolution of the question $i$, 0 otherwise. If all involved skills are necessary to succeed the corresponding item, the model is considered part of the \emph{conjunctive} class. For example, in the following q-matrix, skills 1 and 2 are required to answer the first question, skills 1 and 3 are required for the second question and mastering skill 3 is sufficient to solve last question.
\[ Q = \left(\begin{array}{lll}
1 & 1 & 0\\
1 & 0 & 1\\
0 & 0 & 1
\end{array}\right) \]
If the mastery of one single skill is sufficient to succeed the item, it will be considered part of the \emph{compensatory} class. We consider here the NIDA conjunctive model~\cite{Desmarais2012}: the probability of a correct response at item $i$ is $1 - s_i$ if all skills involved are mastered, $g_i$ if any skill is not mastered. $s_i$ and $g_i$ are respectively called the \emph{slip} and \emph{guess} parameters of item $i$.

Determining the best q-matrix that fits a given student data set is currently an open field of research, as well as the best number of attributes $K$~\cite{Huebner2010} \alert{(maladroit)}, the most known methods being non-negative matrix factorization or EM algorithm~\cite{Huebner2010}. 

In order to compute the state of an examinee effectively, the method is similar to the IRT model: we maintain a probability distribution over the $2^K$ possible states~\cite{Huebner2010} and each question answered allows to update this distribution using the Bayes rule. Several methods for next item selection have been compared, such as Shannon entropy~\cite{Xu2003}.

From this probability distribution over all possible states, we can get the probability for our examinee to get skill $j$ by summing the probabilities of all states having the $j$-th bit at 1. Thus, we can derive the probability for her to answer correctly every question in the bank, with the help of our q-matrix.

\begin{table}
\caption{Example for $K = 3$.}
\end{table}

\section{Our Test Framework}

Our student data is a binary matrix students $\times$ questions where $c_{ij}$ is 1 if student $i$ answered the question $j$ correctly, 0 otherwise. 

We used the following cross-validation based method: 

\begin{itemize}
\item \textsc{TrainingStep} : Train the model with all the student but one in order to produces questions parameters. In the case of an IRT based model questions difficulties are computed, in the case of CDM based model this is the q-matrix which is computed. 
\item \textsc{PriorInitialisation} : Initialize the prior on the student parameters.  These parameters are the distribution on the skills in the case of CDM based model and the distribution on the ability for IRT based model. 
\item \textsc{NextItem} : Choose the best question to ask according to a certain criterion given all previous questions and answers. 
\item \textsc{UpdateParameters} : Given the question just asked and the answer given, we update the student parameters.
\item \textsc{TerminationRule} : "There have been $B$ questions asked so far". Indeed, as we want to compare different algorithm, we need them on an equal footing (i.e. the same number of questions asked). 
\item \textsc{PredictPerformance} : Predict answers of the student on the whole test.
\item \textsc{Evaluate} : Compare the true answers with the predicted ones in order to produce a performance indicator of the algorithm. 
\end{itemize}

\begin{algorithm}
\caption*{\textbf{Computerized Adaptive Testing Framework}}
\begin{algorithmic}
\Procedure{Test}{}
\For{all students s}
	\State $\alpha \gets \Call{TrainingStep}{\text{all students except s}}$
	\State $\beta \gets \Call{PriorInitialisation}$
	\State $i \gets 0$
	\While{\textsc{TerminationRule} is not satisfied}
		\State $q_{i + 1} \gets \Call{NextItem}{q_1, r_1, \ldots, q_i, r_i, \alpha, \beta}$
		\State Ask question $q_{i + 1}$ to the examinee $s$
		\State Get reply $r_{i + 1}$
		\State $\beta \gets \Call{EstimateParameters}{q_1, r_1, \ldots, q_i, r_i, \alpha}$
	\EndWhile
	\State $p \gets$ \Call{PredictPerformance}{$\alpha, \beta$}
	\State \Call{Evaluate}{$p$}
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Item Response Theory Design}

We used the R packages ltm~\cite{Rizopoulos2006} and catR~\cite{MagisRaiche2012} for this part of our experiments.

\subsubsection{Training Step}

We compute for each question $j$ its difficulty $d_j$ along with its discrimination $\delta_j$ and also for each student $i$ her ability $\theta_i$ and the variance associated $v_i$ which satisfied the maximum likelihood. Let $x_i$ being the unobserved true ability of students, the log-likelihood can be computed using Baye formula: 

\[\log \mathcal{L} = \sum_{i,j} \log \int \frac{P(x_i | \theta_i, v_i)dx_i}{1+\exp(\delta_j (d_j - \theta))}\]

There are different way of maximizing the log-likelihood and the R package uses a variant of the E-M algorithm. 

\subsubsection{Item Selection Rule}

The R package minimize the expected variance of the posterior distribution (MEPV criterion) which is actually equivalent as minimizing the differential Shannon entropy in the normal and logistic case. Indeed, these two distribution present a linear link between the logarithm of the variance and the Shannon entropy. 

\subsubsection{Parameters Estimation}

After each student's answer, we can recompute the posterior distribution over the student ability. The R package uses a variant of the E-M algorithm. 

Note that for a normal or a logistic prior, the posterior distribution is no longer normal or logistic. Therefore for an exact computation of the expectation and variance of the student ability, all the previous questions and answers are needed for each update. 

\subsection{Cognitive Diagnosis Model Design}

\subsubsection{Training Step}

We want to compute the q-matrix along with a guess and a slip parameters which minimize the likelihood. The computation of q-matrix is still an open field of research. Recent work of Desmarais \alert{cite} makes use of Non-negative Matrix Factorization. \alert{to be continued}

\subsubsection{Item Selection Rule}

As an item selection rule, we choose the question minimizing the Shannon entropy of the distribution over the possible states. Please note that this is the same logic that the MEPV criterion used in the IRT design~\cite{Cheng2009}.

\subsubsection{Parameters Estimation}

As the distribution of skill is on a finite support, we can at each step maintain the whole probability distribution $\pi$ over the possible states, initialized at the uniform distribution.
Knowing the student answer, the update of $\pi$ is done according to the Bayes rule. Let $x$ be skills, $s$ the question slip and $g$ the question guessing, if the skills $x$ are enough to answer correctly to the question
\[ \pi_{n+1}(x) = \pi_n(x) \cdot [r\cdot(1-s) + (1-r)\cdot s] \]
otherwise
\[ \pi_{n+1}(x) = \pi_n(x) \cdot [r\cdot g + (1-r)\cdot(1-g)] \]

\subsection{Our Evaluation Framework}

We choose the log-likelihood to compare both performances.

\section{Results}

\subsection{Simulated data}

Here we first generate a random q-matrix and guessing and slipping parameters, then we determine the student data accordingly. Therefore our q-matrix is perfectly calibrated.

\subsection{True data}

LSAT.

\alert{Peut-être mettre plus que deux algorithmes (IRT et q-matrix) sur le graphe ?}

\section{Discussion and Future Work}

Add a last column to q-matrix

No study has been done yet over the number of skills

\section{Conclusions}

Outperforms on both simulated data and real data.

%\section{Acknowledgments}
%
%We wish to thank JBG for his valuable comments.

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
%Appendix A
%\section{If I have something more to say}

%\balancecolumns % GM June 2007
% That's all folks!
\end{document}
